---
title: "Final Project"
author: "Aaron Guerra"
format: html
---

The goal of this project is to compare supervised and unsupervised methods for topic modelling using the existing manual codes as gold standard data. To do supervised topic modelling, I train a random forest classifier to predict if a certain topic is being discussed in a policy based on the text of that policy. To do unsupervised topic modelling, I use LDA to identify clusters of words that are commonly used together across policies, and then I interpret those clusters as topics.

# From Homework 1

First, I load in the data directly from the github repo where it is stored and do some initial cleaning so that I can better look at the topics of interest. To start, I plot the number of policies per city and the number of times each topic is considered across all cities.

```{r}

library(readxl)
library(tidyverse)
library(quanteda)
library(randomForest)

# download data from Bay Adapt Github
url <- "https://github.com/BCDC-GIS/Bay-Adapt-Currents/blob/e8e4ec0ad1ee6888070975a138a3d7d075c1a40f/Sea%20Level%20Rise%20in%20General%20Plans/Cities.xlsx?raw=true"

# download.file(url, destfile = "data/Cities_Simplified.xlsx")

cities_data <- read_excel("data/Cities_Simplified.xlsx")[,1:41]

colnames(cities_data) <- c(
  'city', 'county', 'doc_type', 'year', 'element', 'policy', 'program', 'text',
  'aff_housing', 'beach', 'building_code', 'community_engage', 'contamination',
  'coordination', 'cult_hist_pres', 'development', 'econ_jobs', 'education_outreach',
  'emergency', 'erosion', 'expenditures', 'funding', 'planning', 'habitat_env_qual',
  'infrastructure', 'managed_retreat', 'mapping', 'levees_seawall', 'nature_based',
  'open_space', 'public_health', 'research_monitor', 'training', 'transport_mobility',
  'utilities', 'vulnerable_populations', 'water_quality', 'wetlands',
  'flood_stormwater', 'groundwater', 'slr_mentioned'
)


# 1. Built Environment
vars_built <- c('infrastructure', 'transport_mobility', 'utilities', 
                'building_code', 'development', 'levees_seawall', 'flood_stormwater')

# 2. Natural Resources
vars_nature <- c('wetlands', 'nature_based', 'habitat_env_qual', 
                 'open_space', 'beach', 'erosion', 'water_quality', 
                 'groundwater', 'contamination')

# 3. Social Equity
vars_social <- c('vulnerable_populations', 'aff_housing', 'public_health', 
                 'cult_hist_pres', 'econ_jobs', 'community_engage', 'education_outreach')

# 4. Governance & Process
vars_gov <- c('planning', 'coordination', 'funding', 'expenditures', 
              'research_monitor', 'mapping', 'training', 'emergency', 'managed_retreat')

cities_clean <- cities_data |>
  mutate(across(aff_housing:slr_mentioned, 
                ~ case_when(. == "1" ~ 1, TRUE ~ 0))) |>
  mutate(year = as.numeric(as.character(year))) |>
  rowwise() |>
  mutate(
    sys_built   = sum(c_across(all_of(vars_built))),
    sys_nature  = sum(c_across(all_of(vars_nature))),
    sys_social  = sum(c_across(all_of(vars_social))),
    sys_gov     = sum(c_across(all_of(vars_gov))),
  ) |>
  mutate(
    sys_built_b = as.factor(ifelse(sys_built > 0, 1, 0)),
    sys_nature_b = as.factor(ifelse(sys_nature > 0, 1, 0)),
    sys_social_b = as.factor(ifelse(sys_social > 0, 1, 0)),
    sys_gov_b = as.factor(ifelse(sys_gov > 0, 1, 0))) |> 
  ungroup() |>
  mutate(across(aff_housing:slr_mentioned, as.factor)) |>
  mutate(across(city:element, as.factor)) |> 
  select(city, element, sys_built_b, sys_nature_b, sys_social_b, sys_gov_b, year, 
         text, county, doc_type, policy, program)
  

```


```{r}
# cities and how many policies they have
cd_cities_plot <- count(cities_data, city) |> 
  arrange(desc(n)) 

# plot
policies_per_city <- ggplot(cd_cities_plot, aes(x = reorder(city, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Policies per City ",
       x = "City",
       y = "Number of Policies") +
  theme_minimal()

#save as 6 by 10 in image
ggsave("images/policies_per_city.png", bg='white', width = 6, height = 10)
```


```{r}
# count of how many times each topic considered
cd_topic_sums <- cities_clean |> 
  select(sys_built_b:sys_gov_b) |> 
  mutate(across(everything(), as.numeric)) |>
  summarise(across(everything(), sum, na.rm = TRUE)) |> 
  pivot_longer(everything(), names_to = "topic", values_to = "count") |> 
  arrange(desc(count))

# plot
ggplot(cd_topic_sums, aes(x = reorder(topic, count), y = count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Count of Topics",
       x = "Topic",
       y = "Count") +
  theme_minimal()


```

In this step, I try to predict if a certain policy is discussing a certain topic on the basis of the text in that policy. To do so I use `quanteda` to create a document-feature matrix from the text of each policy, such that each row is a policy and each column is a word, thus, p>>n. However, I filter to only the top 1000 words which appear in the dataset, which reduces p<n. Then I train an RF classifier to predict if a policy discusses the topic (3 are used here as examples) based on the words used in the policy. 

```{r}

cities_text <- cities_clean %>%
  pull(text)

slr_docvar <- cities_clean %>%
  select(sys_built_b:sys_gov_b)

# convert cities_data$text to a corpus
cities_dfm <- corpus(cities_clean, docvars=slr_docvar) %>% 
  tokens(remove_punct = TRUE) %>%
  tokens_remove(stopwords("en")) %>% 
  dfm() %>% 
  dfm_trim(min_termfreq = 500,
           termfreq_type = "rank")

set.seed(354651)
train_index <- sample(1:nrow(cities_clean), 
                      size = 0.7 * nrow(cities_clean), 
                      replace = FALSE)

cities_dfm_train <- cities_dfm[train_index, ]

dfm_df_train <- convert(cities_dfm, to='data.frame') %>% 
  select(-doc_id) |> 
  tibble() |> 
  slice(train_index)

built_rf <- randomForest(x = dfm_df_train, y = docvars(cities_dfm_train, "sys_built_b"), ntree = 500); built_rf
nature_rf <- randomForest(x = dfm_df_train, y = docvars(cities_dfm_train, "sys_nature_b"), ntree = 500); nature_rf
social_rf <- randomForest(x = dfm_df_train, y = docvars(cities_dfm_train, "sys_social_b"), ntree = 500); social_rf
gov_rf <- randomForest(x = dfm_df_train, y = docvars(cities_dfm_train, "sys_gov_b"), ntree = 500); gov_rf

varImpPlot(built_rf)
varImpPlot(nature_rf)
varImpPlot(social_rf)
varImpPlot(gov_rf)
```

```{r}
# testing on test set
dfm_df_test <- convert(cities_dfm, to='data.frame') %>% 
  select(-doc_id) |> 
  tibble() |> 
  slice(-train_index)


cities_dfm_test <- cities_dfm[-train_index, ]


built_pred <- predict(built_rf, newdata = dfm_df_test)
built_score <- mean(built_pred == docvars(cities_dfm_test, "sys_built_b")); 1-built_score

nature_pred <- predict(nature_rf, newdata = dfm_df_test)
nature_score <- mean(nature_pred == docvars(cities_dfm_test, "sys_nature_b")); 1-nature_score

social_pred <- predict(social_rf, newdata = dfm_df_test)
social_score <- mean(social_pred == docvars(cities_dfm_test, "sys_social_b")); 1-social_score

gov_pred <- predict(gov_rf, newdata = dfm_df_test)
gov_score <- mean(gov_pred == docvars(cities_dfm_test, "sys_gov_b")); 1-gov_score


```

Now to do unsupervised topic modelling, I use Seeded LDA to identify words that belong together 

```{r}
library(seededlda)

set.seed(354651)
lda_model <- textmodel_lda(cities_dfm_train, k=4)

terms(lda_model, 10)


```

Based on this it seems like a fairly decent start. The first topic seems to look at built infrastructure, with terms like "flood", "facilities", and "development". The third topic maps fairly naturally to natural resources with terms like "shoreline", "habitat", "creek", and "natural". The thid topic seems to get at governance, with terms like "city", "mitigation", "plan", and "planning". The fourth topic sort of misses the mark, however, and does not get at social equity and seems to be a more water-focused version of built infrastructure. This is somewhat to be expected with LDA, as only four topics will be pretty coarse by nature and will not cover everything expected. There are two directions to go from here - one is to fit another completely unsupervised LDA model with more categories which we can aggregate and look through to get at a combination of topics that map to social equity.

The other option is to use some of the information we have, namely, that each topic contains certain subtopics. We can use the terms in these subtopics to seed the LDA model, which will hopefully give four topics that match what we are looking for. This is a bit of a rough approach without *too* much thought, where I just take the main words from each subtopic and use them. A more complicated version would start to concatenate phrases that appear together frequently, for example, instead of "sea", "level", and "rise" being seperate tokens, we might concatenate them to "sea level rise". However, this should suffice for this project. 

```{r}

# 1. Built Environment
vars_built <- c('infrastructure', 'transport_mobility', 'utilities', 
                'building_code', 'development', 'levees_seawall', 'flood_stormwater')

# 2. Natural Resources
vars_nature <- c('wetlands', 'nature_based', 'habitat_env_qual', 
                 'open_space', 'beach', 'erosion', 'water_quality', 
                 'groundwater', 'contamination')

# 3. Social Equity
vars_social <- c('vulnerable_populations', 'aff_housing', 'public_health', 
                 'cult_hist_pres', 'econ_jobs', 'community_engage', 'education_outreach')

# 4. Governance & Process
vars_gov <- c('planning', 'coordination', 'funding', 'expenditures', 
              'research_monitor', 'mapping', 'training', 'emergency', 'managed_retreat')

seed_dict <- dictionary(list(
  built = c('infrastructure', 'transport*', 'mobility', 'utilit*', 'building*', 
           'development', 'levees', 'seawall', 'flood*','stormwater*'),
  nature = c('wetland*', 'natur*', 'habitat*', 'environmental', 'quality', 
             'space', 'beach*', 'erosion', 'groundwater', 'contamination'),
  social = c('vulnerab*','population*', 'affordab*', 'health', 
             'cultur*', 'histor*', 'econom*', 'job*', 'communit*', 
             'engagement', 'education*', 'outreach'),
  gov = c('plan*', 'coordinat*', 'fund*', 'expenditur*', 
          'research*', 'monitor', 'map*', 'train*', 'emergenc*')
))

set.seed(354651)
seeded_lda_model <- textmodel_seededlda(cities_dfm_train, 
                                         dictionary=seed_dict)
terms(seeded_lda_model, 10)
```

So based on these four topics, it seems that the fourth topic that was previously hard to decipher is mostly mapping to the 'social equity' topic in the seeded model. The 'built' and 'nature' topics remain fairly similar, and the 'governance' topic seems to shift a little bit. There is certainly room for iterative, mostly manual improvement here, in refining the dictionary and evaluating some of these terms in context (all of which can be done in the `quanteda` package). 

```{r}
# topic 1 - infrasturcture
# topic 2 - nature
# topic 3 - governance
# topic 4 - social equity

pred_lda <- textmodel_lda(cities_dfm_test, 
                             model=lda_model)

built_pred_lda <- as.factor(as.numeric(topics(pred_lda) == 'topic1'))
built_score_lda <- mean(built_pred_lda == docvars(cities_dfm_test, "sys_built_b")); 1-built_score_lda

nature_pred_lda <- as.factor(as.numeric(topics(pred_lda) == 'topic2'))
nature_score_lda <- mean(nature_pred_lda == docvars(cities_dfm_test, "sys_nature_b")); 1-nature_score_lda

social_pred_lda <- as.factor(as.numeric(topics(pred_lda) == 'topic4'))
social_score_lda <- mean(social_pred_lda == docvars(cities_dfm_test, "sys_social_b")); 1-social_score_lda

gov_pred_lda <- as.factor(as.numeric(topics(pred_lda) == 'topic3'))
gov_score_lda <- mean(gov_pred_lda == docvars(cities_dfm_test, "sys_gov_b")); 1-gov_score_lda

```

```{r}
# topic 1 - infrastructure
# topic 2 - nature
# topic 3 - governance
# topic 4 - social equity

pred_seeded <- textmodel_lda(cities_dfm_test, 
                             model=seeded_lda_model)

built_pred_lda_seed <- as.factor(as.numeric(topics(pred_seeded) == 'built'))
built_score_lda_seed <- mean(built_pred_lda_seed == docvars(cities_dfm_test, "sys_built_b")); 1-built_score_lda_seed

nature_pred_lda_seed <- as.factor(as.numeric(topics(pred_seeded) == 'nature'))
nature_score_lda_seed <- mean(nature_pred_lda_seed == docvars(cities_dfm_test, "sys_nature_b")); 1-nature_score_lda_seed

social_pred_lda_seed <- as.factor(as.numeric(topics(pred_seeded) == 'social'))
social_score_lda_seed <- mean(social_pred_lda_seed == docvars(cities_dfm_test, "sys_social_b")); 1-social_score_lda_seed

gov_pred_lda_seed <- as.factor(as.numeric(topics(pred_seeded) == 'gov'))
gov_score_lda_seed <- mean(gov_pred_lda_seed == docvars(cities_dfm_test, "sys_gov_b")); 1-gov_score_lda_seed

```


